TAPIR: Threaded Application Programming on Intermittent Resources

### 2017-05-12 paper.tex
Cleaned up and marked TODOs
abstract

%Only including an abstract since it looks like the submission site wants one...
\begin{abstract}
Killer abstract.
\end{abstract}

@@

TAPIR has a round robin scheduling algorithm. At the end of a task, the compiler
inserts a jump back to the scheduler task, where the next task in the list is
the one to be executed. An alternate scheduling model that was considered was
a context switch on reboot instead of at the end of each task. The main drawback
of using a reboot-based scheduler is that, because power failure is the common
case, there is a high chance of reducing or halting forward progress altogether
because one thread holds a lock that all of the other threads need but cannot
complete. For that reason the design specifies end of task, so each task has
an opportunity to use a shared resource and release any locks before a power-
intensive context switch occurs. Additionally, a round robin scheduler was
chosen for this design because it is simple to implement and uses only one index counter,
very little power or memory overhead as compared to other scheduling algorithms. It is
critical for multithreading in Chain to not add significant power and memory
overhead in scheduling over conventional Chain so that the workload can
make as much forward progress as possible on one charge cycle.

>>>>

%Only including an abstract since it looks like the submission site wants one...
\begin{abstract}
The growth of research in IoT related wireless technologies and low-power computing
has led to the new area of \textit{energy-harvesting computation} on \textit{intermittently
powered devices}. Intermittently powered devices are governed by the \textit{intermittent
execution model} where regular reboots due to an energy charge cycle necessitate division
of a program into small idempotent \textit{tasks}. However, current applications for
intermittently powered devices are small and generally contain little functionality.

In this paper we introduce the \textit{Threaded Application Programming on Intermittent Resources} 
(TAPIR) model to expand the current intermittent programming model \textit{Chain} to 
include general multiprogramming. TAPIR allows the range of possible programs for 
intermittent devices to include more general computation functionality while leaving the
memory and energy overhead low enough that intermittent computation on those workloads 
still guarantees forward progress. While nonvolatile memory usage will increase per thread
scheduled, certain optimizations are possible to reduce memory overhead overall. Additionally,
context switching through our scheduler task is a low-instruction count static overhead that is
easily accounted for in energy calculations. Through the design of TAPIR we show that general
purpose threaded applications are feasible on an intermittently powered device.

\end{abstract}

@@

TAPIR has a round robin scheduling algorithm. At the end of a task, the compiler
inserts a jump back to the scheduler task, where the next task in the list is
the one to be executed. An alternate scheduling model that was considered was
a context switch on reboot instead of at the end of each task. The main drawback
of using a reboot-based scheduler is that, because power failure is the common
case, there is a high chance of reducing or halting forward progress altogether
because one thread holds a lock that all of the other threads need but cannot
complete. For that reason the design specifies end of task, so each task has
an opportunity to use a shared resource and release any locks before a power-
intensive context switch occurs. Additionally, a round robin scheduler was
chosen for this design because it is simple to implement and uses only one index counter,
very little power or memory overhead as compared to other scheduling algorithms. It is
critical for multithreading in Chain to not add significant power and memory
overhead in scheduling over conventional Chain so that the workload can
make as much forward progress as possible on one charge cycle. While out of
scope for this project, future work may include a priority scheduler to decrease
latency for certain threads due to context switching for I/O or other input latency
sensive workloads.
